Effects of Water Permeability on the Hydrodynamics of a Vesicle Under Flow and Confinement

### 2020-12-28 vesicle_code/docs/permeableVes/main.tex
worked on supp material

\section*{Supplementary Material}
\subsection*{Formulation}
We consider a two-dimensional inextensible semi-permeable vesicle,
$\gamma$, suspended in unconfined and confined geometries. The
suspending fluid is assumed to be viscous so that the fluid velocity
$\uu$ and pressure $p$ are governed by the incompressible Stokes
equations
\begin{align}
  \label{eqn:governing}
  -\nabla p + \mu \triangle \uu = 0, \quad
  \nabla \cdot \uu = 0, \qquad \xx \in \RR^2 \setminus \gamma,
\end{align}
where $\mu$ is the fluid viscosity. In an unconfined geometry, the
condition $\uu(\xx) \rightarrow \uu_\infty(\xx)$ is enforced as $\xx
\rightarrow \infty$, and in a confined geometry with boundary $\Gamma$,
the boundary condition $\uu(\xx) = \UU(\xx)$ is enforced on $\Gamma$,
where $\UU(\xx)$ is prescribed. Along the vesicle $\gamma$, mass
continuity and force balance are enforced by requiring
\begin{align}
  \label{eqn:jump}
  \jump{T\cdot\nn} = \ff_\mathrm{mem}, \quad \jump{\uu} = 0,
\end{align}
where $\nn$ is the outward normal of $\gamma$, $\jump{\cdot}$ is the
jump across the interface, and $T$ is the hydrodynamic stress tensor.
The membrane force $\ff_\mathrm{mem}$ is the sum of a bending force
with bending modulus $k_b$ and a force due to the tension $\Lambda$.
Following~\cite{vee-gue-zor-bir2009}, we modify the bending and tension
forces as $\ff_\mathrm{ben} = -k_b \xx_{ssss}$, and $\ff_\mathrm{ten} =
(\sigma \xx_s)_s$, where $s$ is arclength. The physical tension can be
recovered using the identity $\Lambda = \sigma + 1.5 \kappa^2$, where
$\kappa$ is the membrane curvature.
\todo[inline]{need inextensibility}

Semi-permeability is introduced by modifying the vesicle velocity from
the standard choice. Instead of using the no-slip boundary condition, we
assume the fluid flux through the boundary is proportional to the
mechanical force~\cite{yao-mor2017}. That is, 
\begin{align}
  \label{eqn:vesicleFlux}
  \uu - \dot{\xx} = - \beta (\ff_\mathrm{mem} \cdot \nn) \nn, \qquad
  \xx \in \gamma,
\end{align}
where $\beta$ is the semi-permeability rate.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Dimensionless equations}
The inextensibility condition guarantees that the vesicle's length, $L$,
is conserved, while its area, $A$, can change due to semi-permeability.
Therefore, the reduced area $\alpha = 4\pi A/L^2 \in (0,1]$ is not
conserved. We non-dimensionalize~\eqref{eqn:vesVelocity} by introducing
the length scale $R_0 = L/2\pi$ and defining the time scale $\tau = \mu
R_0^3/k_b$, the tension scale $Q = k_b/R_0^2$, and the permeability
scale $W = R_0/\mu$. The background velocity $\uu_\infty$ is
non-dimensionalized with $R_0/\tau$.

\todo[inline]{This is a dump of information regarding how we
non-dimensionalize}
In each experiment, we initialize an elliptical vesicle with longest
diameter of 9$\mu$m. We use a length scale of $R_0 = 10^{-6}$m, a
bending stiffness of $k_b = 10^{-19}$J, and fluid viscosity of $\mu=5
\times 10^{-2}$kg/ms. Using the relaxation time scale $\tau = \mu
R_0^3/k_b = 5 \times 10^{-1}$s, the pressure scale is $P = k_b/R_0^3 =
10^{-1}$Pa.

>>>>

\section*{Supplementary Material}
\subsection*{Formulation}
We consider a two-dimensional inextensible semi-permeable vesicle,
$\gamma$, suspended in unconfined and confined geometries. The
suspending fluid is assumed to be viscous so that the fluid velocity
$\uu$ and pressure $p$ are governed by the incompressible Stokes
equations
\begin{align}
  \label{eqn:governing}
  -\nabla p + \mu \triangle \uu = 0, \quad
  \nabla \cdot \uu = 0, \qquad \xx \in \RR^2 \setminus \gamma,
\end{align}
where $\mu$ is the fluid viscosity. In an unconfined geometry, the
condition $\uu(\xx) \rightarrow \uu_\infty(\xx)$ is enforced as $\xx
\rightarrow \infty$, and in a confined geometry with boundary $\Gamma$,
the boundary condition $\uu(\xx) = \UU(\xx)$ is enforced on $\Gamma$,
where $\UU(\xx)$ is prescribed. Along the vesicle $\gamma$, mass
continuity, a force balance, and local inextensibility are enforced by
requiring
\begin{align}
  \label{eqn:jump}
  \jump{\uu} = 0, \quad
  \jump{T\cdot\nn} = \ff_\mathrm{mem}, \quad
  \xx_s \cdot \uu_s = 0,
\end{align}
where $\nn$ is the outward normal of $\gamma$, $\jump{\cdot}$ is the
jump across the interface, and $T$ is the hydrodynamic stress tensor.
The membrane force $\ff_\mathrm{mem}$ is the sum of a bending force
with bending modulus $k_b$ and a force due to the tension $\Lambda$.
Following~\cite{vee-gue-zor-bir2009}, we modify the bending and tension
forces as $\ff_\mathrm{ben} = -k_b \xx_{ssss}$, and $\ff_\mathrm{ten} =
(\sigma \xx_s)_s$, where $s$ is arclength. The physical tension can be
recovered using the identity $\Lambda = \sigma + 1.5 \kappa^2$, where
$\kappa$ is the membrane curvature.

Semi-permeability is introduced by modifying the vesicle velocity from
the standard choice. Instead of using the no-slip boundary condition, we
assume the fluid flux through the boundary is proportional to the
mechanical force~\cite{yao-mor2017}. That is, 
\begin{align}
  \label{eqn:vesicleFlux}
  \uu - \dot{\xx} = - k_w (\ff_\mathrm{mem} \cdot \nn) \nn, \qquad
  \xx \in \gamma,
\end{align}
where $k_w$ is the water permeability.

The governing equations are non-dimensionalized by introducing a length
scale that we take to be $R_0 = 10^{-6}$m. Using a bending stiffness of
$k_b = 10^{-19}$J and fluid viscosity $5 \times 10^{-2}$kg/ms, we define
the time scale $\tau = \mu R_0^3/k_b = 0.5s$, the velocity scale $U =
R_0/\tau = 2\mu$m/s, the pressure scale $P = k_b/R_0^3 = 10^{-1}$Pa, the
tension scale $Q = k_b/R_0 = 10^{-13}$N, and the permeability scale $W =
R_0\mu = 2 \times 10^{-5}$m$^2$s/kg. The other standard dimensionless
variable of a vesicle is its reduced area $\nu = 4\pi A/L^2 \in (0,1]$.
The length $L$ is conserved by the local inextensibility condition.
However, the area $A$ is not conserved because of semi-permeability, and
therefore $\nu$ is also not conserved.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection*{Dimensionless equations}
%The inextensibility condition guarantees that the vesicle's length, $L$,
%is conserved, while its area, $A$, can change due to semi-permeability.
%Therefore, the reduced area $\alpha = 4\pi A/L^2 \in (0,1]$ is not
%conserved. We non-dimensionalize~\eqref{eqn:vesVelocity} by introducing
%the length scale $R_0 = L/2\pi$ and defining the time scale $\tau = \mu
%R_0^3/k_b$, the tension scale $Q = k_b/R_0^2$, and the permeability
%scale $W = R_0/\mu$. The background velocity $\uu_\infty$ is
%non-dimensionalized with $R_0/\tau$.
%
%In each experiment, we initialize an elliptical vesicle with longest
%diameter of 9$\mu$m. We use a length scale of $R_0 = 10^{-6}$m, a
%bending stiffness of $k_b = 10^{-19}$J, and fluid viscosity of $\mu=5
%\times 10^{-2}$kg/ms. Using the relaxation time scale $\tau = \mu
%R_0^3/k_b = 5 \times 10^{-1}$s, the pressure scale is $P = k_b/R_0^3 =
%10^{-1}$Pa.
